ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel

### 2025-01-22 paper/Manuscript-Data-Package.tex
Some cosmetic fixes to paper

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{general-social-survey-gss-collection}{%
\section{General Social Survey (GSS)
collection}\label{general-social-survey-gss-collection}}

@@

\hypertarget{the-main-file}{%
\subsection{The Main File}\label{the-main-file}}

@@

\hypertarget{the-episode-file}{%
\subsection{The Episode File}\label{the-episode-file}}

@@

\hypertarget{data-extraction}{%
\section{Data extraction}\label{data-extraction}}

@@

\hypertarget{how-to-use-activeca}{%
\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}}

@@

By running \texttt{?ActiveCA} in \texttt{R}, you can access the complete
package documentation. The documentation includes a description of the
package, information about the authors and maintainers, and an overview
of all pre-processed data sets included in the package.

@@

The chunk below shows how to load the walking episodes from the 2015 GSS
(Cycle 29):

@@

The GSS surveys apply a probability sampling methodology, in which each
episode or person selected in the sample represents several other
episodes or persons not in the sample. The number of episodes and
persons represented by a episode or person is determined by the weight
or weighting factor. Because of this, every estimates of the number of
episodes or persons need to be calculated applying the corresponding
weighting factors.

For instance, if someone wants to calculate the percentage of
respondents from the 2015 GSS survey with active travel episodes, it is
necessary to account for the person's weights. In 2015, the weight
variable is represented by \texttt{WGHT\_PER.} The code chunk below
demonstrates how to obtain the percentage of people with active travel
episodes by age group. It uses the \texttt{dplyr}package
\citep{dplyr2023} to manipulate the data.

@@

# Calculate the total population for each age group
Total_population <- gss_main_2015 %>% 
  group_by(AGEGR10) %>% 
  summarise(Total_population  = sum(WGHT_PER))

# Calculate the active population for each age group
Active_population <- gss_episodes %>% 
      filter(YEAR == 2015) %>% 
  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) %>% 
  group_by(AGEGR10) %>% 
  summarise(Active_population  = sum(WGHT_PER))

# Combine total and active population 
#and calculate the percentage
Active_percentage <- Total_population %>% 
  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
  mutate(
    Percentage = round(Active_population/Total_population*100,2))

# Display the final result
Active_percentage
\end{verbatim}

\begin{verbatim}
# A tibble: 7 x 4
  AGEGR10           Total_population Active_population Percentage
  <ord>                        <dbl>             <dbl>      <dbl>
1 15 to 24 years            4511131.          1641468.       36.4
2 25 to 34 years            4956386.          1644033.       33.2
3 35 to 44 years            4734506.           847770.       17.9
4 45 to 54 years            5136125.           790667.       15.4
5 55 to 64 years            4831306.           767064.       15.9
6 65 to 74 years            3283969.           559025.       17.0
7 75 years and over         2312976.           384359.       16.6
\end{verbatim}

\hypertarget{processed-data-set}{%
\subsection{Processed data set}\label{processed-data-set}}

@@

\hypertarget{descriptive-statistics}{%
\subsection{Descriptive statistics}\label{descriptive-statistics}}

@@

\hypertarget{python-integration}{%
\section{Python integration}\label{python-integration}}

@@

\hypertarget{concluding-remarks}{%
\section{Concluding remarks}\label{concluding-remarks}}

@@

>>>>

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

@@

\section{Introduction}\label{introduction}

@@

\section{General Social Survey (GSS)
collection}\label{general-social-survey-gss-collection}

@@

\subsection{The Main File}\label{the-main-file}

@@

\subsection{The Episode File}\label{the-episode-file}

@@

\section{Data extraction}\label{data-extraction}

@@

\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}

@@

The documentation of the package can be accessed by running
\texttt{?ActiveCA} in \texttt{R}. The documentation includes a
description of the package, information about the authors, contributors,
and maintainers, and an overview of all pre-processed data sets included
in the package.

@@

The code below shows how to load the walking episodes from the 2015 GSS
(Cycle 29):

@@

The GSS surveys apply a probability sampling methodology, in which each
episode or person selected in the sample represents several other
episodes or persons not in the sample. The number of episodes and
persons represented by a episode or person is determined by the weight
or weighting factor. Because of this, estimates of the number of
episodes or persons need to be calculated applying the corresponding
weighting factors.

For instance, to calculate the percentage of respondents from the 2015
GSS survey with active travel episodes, it is necessary to account for
the person's weight. In 2015, the weight variable is represented by
\texttt{WGHT\_PER.} The code below demonstrates how to obtain the
percentage of people with active travel episodes by age group. It uses
the \texttt{dplyr} package \citep{dplyr2023} to manipulate the data.

@@

# Calculate the total population for each age group
Total_population <- gss_main_2015 |> 
  group_by(AGEGR10) |> 
  summarise(Total_population  = sum(WGHT_PER))

# Calculate the active population for each age group
Active_population <- gss_episodes |> 
      filter(YEAR == 2015) |> 
  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) |> 
  group_by(AGEGR10) |> 
  summarise(Active_population  = sum(WGHT_PER))

# Combine total and active population 
#and calculate the percentage
Active_percentage <- Total_population |> 
  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
  mutate(
    Percentage = round(Active_population/Total_population*100,2))
\end{verbatim}

\begin{longtable}[]{@{}lrrr@{}}
\toprule\noalign{}
AGEGR10 & Total\_population & Active\_population & Percentage \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
15 to 24 years & 4511131 & 1641468.3 & 36.39 \\
25 to 34 years & 4956386 & 1644032.8 & 33.17 \\
35 to 44 years & 4734506 & 847769.9 & 17.91 \\
45 to 54 years & 5136125 & 790667.3 & 15.39 \\
55 to 64 years & 4831306 & 767064.0 & 15.88 \\
65 to 74 years & 3283969 & 559024.6 & 17.02 \\
75 years and over & 2312976 & 384358.6 & 16.62 \\
\end{longtable}

\subsection{Processed data set}\label{processed-data-set}

@@

\subsection{Descriptive statistics}\label{descriptive-statistics}

@@

\section{Python integration}\label{python-integration}

@@

\section{Concluding remarks}\label{concluding-remarks}

@@

For review purposes, the package is currently available as a tar.gz file
that can be installed by R. The file can be obtained from this anonymous
location:

\url{https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip}
